---
title: "Households With An Improved Drinking Water Source in India (NFHS-4)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rgdal::readOGR
```

##Introduction

"Tough competition for limited water resources, and poor decision-making by governments and utilities on prioritising how those resources are used, are already making it hard for the world’s poorest people to access clean water. Land use alters as populations, agriculture and industry move, change and grow; if it isn’t controlled, the result may be land erosion, pollution and depletion of groundwater. 

Wild water events can wipe out fragile infrastructure, dry up rivers, ponds and springs which are sometimes the main source of water for the poorest people, and contribute to the spread of waterborne diseases. Rural populations in poor and geographically isolated areas face particular challenges. Of all the people in the world without access to clean water, more than half a billion – enough to circle the world over six times – are in rural areas.

Here, help is often slow to arrive after natural disasters, infrastructure is poor to non-existent, and a continued lack of funding is most acutely felt." - [Wild Water] (http://wateraidindia.in/publication/wild-water-state-worlds-water-2017/)

In this analysis I am trying to measure how many regions in India have access to clean drinking water on a state and district level. The darker the colouring gets, the households in that region are more likely to have access to clean drinking water. 

The data is sourced from [National Family Health Survey](http://rchiips.org/nfhs/). 

## R Markdown

```{r libraries, include=FALSE}
library(maps)
library(maptools)
library(rgdal)
```
## Cleaning And Importing Data
```{r datacleaning, cache=FALSE}
data <-read.csv("data.csv")
data$ImprovedDrinkingWater<-data$Households.with.an.improved.drinking.water.source1
data$ImprovedDrinkingWater<-data$ImprovedDrinkingWater/100
data1 <- aggregate(ImprovedDrinkingWater~State.1, data=data, FUN=mean)

datatime <-read.csv("states_improved_water_source_percentages.csv")
datatime$four<-datatime$Percent_with_improved_water...4
datatime$four<-datatime$four/100
datatime$three<-datatime$Percent_with_improved_water...3
datatime$three<-datatime$three/100
```

```{r NFHS-3, echo=FALSE, warning=FALSE}
ind<- readShapePoly("IND_adm/IND_adm1", 
                    proj4string =CRS("+proj=aea +lat_1=28 +lat_2=12 +lat_0=20 +lon_0=78 +x_0=2000000 +y_0=2000000 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
pickColor <- function(x) {
  if (is.na(x)) {
    col <- "#ffffff"
  }
  else if (x < 0.5) {
    col <- "#f0f0f0"
  }
  else if (x < 0.6) {
    col <- "#cecece"
  }
  else if (x < 0.7) {
    col <- "#bdbdbd"
  }
  else if (x < .8) {
    col <- "#a5a5a5"
  }
  else if (x < .9) {
    col <- "#939393"
  }
  else {
    col <- "#636363"
  }
  return(col)
}

datatime$col <- sapply(datatime$three, pickColor)

plot(ind, col=datatime$col, border="#ffffff", lwd=0.5,  main="NFHS-3")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)
```
```{r NFHS-4, echo=FALSE, warning=FALSE}
datatime$col <- sapply(datatime$four, pickColor)

plot(ind, col=datatime$col, border="#ffffff", lwd=0.5,  main="NFHS-4")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)
```

## States And Districts (NFHS-4)

```{r Andaman and Nicobar, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_andamanandnicobar <- ind_district[ind_district$NAME_1 == "Andaman and Nicobar",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_andamanandnicobar, col=data$col, border="#ffffff", lwd=0.5, main = "Andaman and Nicobar")
par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Andhra Pradesh, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_andhrapradesh <- ind_district[ind_district$NAME_1 == "Andhra Pradesh",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_andhrapradesh, col=data$col, border="#ffffff", lwd=0.5, main = "Andhra Pradesh")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Arunachal Pradesh, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_arunachalpradesh <- ind_district[ind_district$NAME_1 == "Arunachal Pradesh",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_arunachalpradesh, col=data$col, border="#ffffff", lwd=0.5, main = "Arunachal Pradesh")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Assam, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_assam <- ind_district[ind_district$NAME_1 == "Assam",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_assam, col=data$col, border="#ffffff", lwd=0.5, main = "Assam")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Bihar, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_bihar <- ind_district[ind_district$NAME_1 == "Bihar",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_bihar, col=data$col, border="#ffffff", lwd=0.5, main = "Bihar")


par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Chandigarh, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_chandigarh <- ind_district[ind_district$NAME_1 == "Chandigarh",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_chandigarh, col=data$col, border="#ffffff", lwd=0.5, main = "Chandigarh")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Chhattisgarh, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_chhattisgarh <- ind_district[ind_district$NAME_1 == "Chhattisgarh",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_chhattisgarh, col=data$col, border="#ffffff", lwd=0.5, main = "Chhattisgarh")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Dadra and Nagar Haveli, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_dadranagar <- ind_district[ind_district$NAME_1 == "Dadra and Nagar Haveli",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_dadranagar, col=data$col, border="#ffffff", lwd=0.5, main = "Dadra and Nagar Haveli")


par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Daman and Diu, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_damandiu <- ind_district[ind_district$NAME_1 == "Daman and Diu",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_damandiu, col=data$col, border="#ffffff", lwd=0.5, main = "Daman and Diu")


par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Delhi, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_delhi <- ind_district[ind_district$NAME_1 == "Delhi",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_delhi, col=data$col, border="#ffffff", lwd=0.5, main = "Delhi")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Goa, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_goa <- ind_district[ind_district$NAME_1 == "Goa",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_goa, col=data$col, border="#ffffff", lwd=0.5, main = "Goa")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Gujarat, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_gujarat <- ind_district[ind_district$NAME_1 == "Gujarat",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_gujarat, col=data$col, border="#ffffff", lwd=0.5, main = "Gujarat")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Haryana, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_haryana <- ind_district[ind_district$NAME_1 == "Haryana",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_haryana, col=data$col, border="#ffffff", lwd=0.5, main = "Haryana")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Himachal Pradesh, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_himachalpradesh <- ind_district[ind_district$NAME_1 == "Himachal Pradesh",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_haryana, col=data$col, border="#ffffff", lwd=0.5, main = "Himachal Pradesh")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Jammu and Kashmir, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_jammukashmir <- ind_district[ind_district$NAME_1 == "Jammu and Kashmir",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_jammukashmir, col=data$col, border="#ffffff", lwd=0.5, main = "Jammu and Kashmir")


par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Jharkhand, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_jharkhand <- ind_district[ind_district$NAME_1 == "Jharkhand",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_jharkhand, col=data$col, border="#ffffff", lwd=0.5, main = "Jharkhand")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Karnataka, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_karnataka <- ind_district[ind_district$NAME_1 == "Karnataka",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_karnataka, col=data$col, border="#ffffff", lwd=0.5, main = "Karnataka")


par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Kerala, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_kerala <- ind_district[ind_district$NAME_1 == "Kerala",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_kerala, col=data$col, border="#ffffff", lwd=0.5, main = "Kerala")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Madhya Pradesh, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_madhyapradesh <- ind_district[ind_district$NAME_1 == "Madhya Pradesh",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_madhyapradesh, col=data$col, border="#ffffff", lwd=0.5, main = "Madhya Pradesh")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Maharashtra, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_maharashtra <- ind_district[ind_district$NAME_1 == "Maharashtra",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_maharashtra, col=data$col, border="#ffffff", lwd=0.5, main = "Maharashtra")


par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Manipur, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_manipur <- ind_district[ind_district$NAME_1 == "Manipur",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_manipur, col=data$col, border="#ffffff", lwd=0.5, main = "Manipur")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Meghalaya, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_meghalaya <- ind_district[ind_district$NAME_1 == "Meghalaya",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_meghalaya, col=data$col, border="#ffffff", lwd=0.5, main = "Meghalaya")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.3, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Mizoram, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_mizoram <- ind_district[ind_district$NAME_1 == "Mizoram",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_mizoram, col=data$col, border="#ffffff", lwd=0.5, main = "Mizoram")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Nagaland, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_nagaland <- ind_district[ind_district$NAME_1 == "Nagaland",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_nagaland, col=data$col, border="#ffffff", lwd=0.5, main = "Nagaland")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Orissa, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_orissa <- ind_district[ind_district$NAME_1 == "Orissa",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_orissa, col=data$col, border="#ffffff", lwd=0.5, main = "Orissa")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Puducherry, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_puducherry <- ind_district[ind_district$NAME_1 == "Puducherry",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_puducherry, col=data$col, border="#ffffff", lwd=0.5, main = "Puducherry")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Punjab, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_punjab <- ind_district[ind_district$NAME_1 == "Punjab",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_punjab, col=data$col, border="#ffffff", lwd=0.5, main = "Punjab")


par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Rajasthan, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_rajasthan <- ind_district[ind_district$NAME_1 == "Rajasthan",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_rajasthan, col=data$col, border="#ffffff", lwd=0.5, main = "Rajasthan")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Sikkim, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_sikkim <- ind_district[ind_district$NAME_1 == "Sikkim",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_sikkim, col=data$col, border="#ffffff", lwd=0.5, main = "Sikkim")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Tamil Nadu, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_tamilnadu <- ind_district[ind_district$NAME_1 == "Tamil Nadu",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_tamilnadu, col=data$col, border="#ffffff", lwd=0.5, main = "Tamil Nadu")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Telangana, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_telangana <- ind_district[ind_district$NAME_1 == "Telangana",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_telangana, col=data$col, border="#ffffff", lwd=0.5, main = "Telangana")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Tripura, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_tripura <- ind_district[ind_district$NAME_1 == "Tripura",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_tripura, col=data$col, border="#ffffff", lwd=0.5, main = "Tripura")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Uttar Pradesh, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_uttarpradesh <- ind_district[ind_district$NAME_1 == "Uttar Pradesh",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_uttarpradesh, col=data$col, border="#ffffff", lwd=0.5, main = "Uttar Pradesh")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r Uttaranchal, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_uttaranchal <- ind_district[ind_district$NAME_1 == "Uttaranchal",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_uttaranchal, col=data$col, border="#ffffff", lwd=0.5, main = "Uttaranchal")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```

```{r West Bengal, echo=FALSE, warning=FALSE}

ind_district<- readShapePoly("IND_adm/IND_adm2")
ind_westbengal <- ind_district[ind_district$NAME_1 == "West Bengal",]
data$col <- sapply(data$ImprovedDrinkingWater, pickColor)
plot(ind_westbengal, col=data$col, border="#ffffff", lwd=0.5, main = "West Bengal")

par(new=TRUE, plt=c(0.1, 0.25, .35, .5))
plot(0:1, 0:1, type="n", xlab="", ylab="", asp=1, axes=FALSE)
col <- c("#ffffff", "#f0f0f0", "#cecece", "#bdbdbd", "#a5a5a5","#939393", "#636363")
breaks <- c(2,3,4,5,6,7)
rect_width <- 1 / length(col)
xleft <- 0:(length(col)-1) * rect_width
ybottom <- rep(.1, length(col))
xright <- 1:length(col) * rect_width
ytop <- rep(.2, length(col))
rect(xleft, ybottom, xright, ytop, col=col, border="#e0e0e0", lwd=.5)
text(.5, .35, "HH with Improved\n Drinking Water Source", cex=.5)
text(1:(length(col)-1) * rect_width, .04, labels = breaks, cex=0.3)

```